\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{geometry}
\geometry{margin=1in}

\title{EC441 Lecture 7 Summary \\ Multiple Access Protocols}
\date{}

\begin{document}

\maketitle

\section{Multiple Access Problem}

A broadcast channel is a shared medium where transmissions from any node are received by all others.
Simultaneous transmissions cause collisions, destroying frames.

MAC (Multiple Access Control) protocols determine:
\begin{itemize}
\item When a node may transmit
\item How collisions are handled
\end{itemize}

An ideal MAC protocol:
\begin{itemize}
\item Gives full rate when one node transmits
\item Shares rate fairly among $M$ nodes ($R/M$)
\item Is decentralized
\item Is simple
\end{itemize}

Three MAC families:
\begin{itemize}
\item Channel partitioning
\item Taking turns
\item Random access
\end{itemize}

\section{Channel Partitioning}

\subsection{TDMA}
Time divided into frames and slots.
Each node gets a fixed slot.

Pros:
\begin{itemize}
\item No collisions
\item Guaranteed rate
\end{itemize}

Cons:
\begin{itemize}
\item Idle slots waste capacity
\end{itemize}

\subsection{FDMA}
Bandwidth divided into frequency bands.

Same tradeoff as TDMA.

\subsection{CDMA}
Nodes transmit simultaneously using orthogonal codes.
Used in 3G cellular and GPS.

Partitioning works best under heavy, predictable load.

\section{Taking Turns}

\subsection{Polling}
Coordinator grants permission to transmit.

Pros:
\begin{itemize}
\item No collisions
\end{itemize}

Cons:
\begin{itemize}
\item Coordinator failure risk
\item Control overhead
\end{itemize}

\subsection{Token Passing}
A token circulates; only the token holder transmits.

Pros:
\begin{itemize}
\item Predictable latency
\end{itemize}

Cons:
\begin{itemize}
\item Token loss
\item Ring failure
\end{itemize}

\section{Random Access — ALOHA}

\subsection{Pure ALOHA}

Transmit immediately. Retransmit after collision.

Collision window: $2T$

Throughput:
\[
S = G e^{-2G}
\]

Maximum throughput:
\[
S_{\max} = \frac{1}{2e} \approx 0.18
\]

Efficiency is low due to collisions.

\subsection{Slotted ALOHA}

Time divided into slots of length $T$.

Collision window reduced to $T$.

Throughput:
\[
S = G e^{-G}
\]

Maximum throughput:
\[
S_{\max} = \frac{1}{e} \approx 0.37
\]

Slotted ALOHA is exactly twice as efficient as pure ALOHA.

\section{CSMA — Carrier Sense Multiple Access}

Key idea: listen before transmitting.

Collisions occur only within propagation delay $\tau$.

Efficiency improves when:
\[
a = \frac{\tau}{T} \text{ is small}
\]

Variants:
\begin{itemize}
\item 1-persistent
\item Non-persistent
\item $p$-persistent
\end{itemize}

\section{CSMA/CD — Collision Detection}

Used in classic Ethernet.

Algorithm:
\begin{enumerate}
\item Sense channel
\item Transmit if idle
\item Monitor while transmitting
\item Abort on collision
\item Binary exponential backoff
\end{enumerate}

Binary exponential backoff:
\[
\text{Wait} \in \{0, \ldots, 2^{k}-1\}
\]

Minimum Ethernet frame size ensures collision detection:
\[
L_{\min} \ge 2 \tau R
\]

Classic Ethernet result:
\[
L_{\min} = 64 \text{ bytes}
\]

CSMA/CD efficiency:
\[
\eta = \frac{1}{1 + 5a}
\]

Modern switched Ethernet is full-duplex, so CSMA/CD is mostly historical.

\section{Wireless — CSMA/CA}

Collision detection fails in wireless because:
\begin{itemize}
\item Self-interference
\item Hidden terminal problem
\end{itemize}

WiFi uses collision avoidance:
\begin{itemize}
\item Random backoff before transmission
\item RTS/CTS reservation (optional)
\item ACK-based reliability
\end{itemize}

\section{Protocol Comparison}

\begin{center}
\begin{tabular}{llll}
\toprule
Family & Protocol & Used In & Mechanism \\
\midrule
Partitioning & TDMA & Cellular, satellite & Time slots \\
Partitioning & FDMA & Radio, cable & Frequency bands \\
Taking turns & Token Ring & Legacy LAN & Token \\
Random access & Pure ALOHA & Packet radio & Transmit freely \\
Random access & Slotted ALOHA & Satellite, LTE & Slot sync \\
Random access & CSMA/CD & Classic Ethernet & Detect collisions \\
Random access & CSMA/CA & WiFi & Avoid collisions \\
\bottomrule
\end{tabular}
\end{center}

\section{Key Numbers}

\begin{itemize}
\item Pure ALOHA max: $\approx 18\%$
\item Slotted ALOHA max: $\approx 37\%$
\item Slotted = 2x Pure ALOHA
\item Ethernet minimum frame: $64$ bytes
\item CSMA/CD efficiency: $\eta = \frac{1}{1 + 5a}$
\end{itemize}

\end{document}